title: "Case Study 1"
author: "Lizzy Sterling and Andy Walch"
date: "October 11, 2017"

### Introduction
In this report, we explore `beers` and `breweries` datasets for the 51 states in the US. The steps and procedures taken in this analysis are stipulated below. We successfully merged the two datasets `Beers` dataset which contains a list of 2410 US craft beers to the `Breweries` dataset containing 558 US breweries.

* [`/Analysis`](https://github.com/lizster/CaseStudy1/tree/master/Analysis) directory is the hub for the R scripts that are used in this analysis. The different R scripts are listed below with a brief description of their functions.
    * [`/analysis/BreweryInformation_files`](https://github.com/lizster/CaseStudy1/tree/master/Analysis/BreweryInformation_files): loads the plots used in the script.
    * [`/analysis/data`](https://github.com/lizster/CaseStudy1/tree/master/Analysis/Data): script loads `Beers` and `Breweries` dataset and performs some basic exploratory data analysis.
    
```{r}
knitr::opts_knit$set(root.dir = '/Users/andy/Desktop/CaseStudy1')
```
```{r}
library(readr)
Beers <- read_csv("~/Downloads/US-Breweries-Data-EDA-master/data/Beers.csv")

breweries <- read.csv("~/Downloads/US-Breweries-Data-EDA-master/data/Breweries.csv")
```

### Research Questions

In this section, we address the research questions put together on this two dataset. The questions are numbered 1 to 7

#### 1. How many breweries are present in each state?

We answer this question by using `count` function in base to count the number of breweries grouped by "state". This prints put a dataframe with two columns `state` which is the State name and `freq` which is the number of breweries in a given state. Each row represents one State. Colorado has the highest number of breweries {47} where as Washington DC, South Dakota (SD), North Dakota (ND), and West Virginia (WV) tie for the least amount of breweries each with just 1.
```{r}
#Andy Code 
count(breweries, "State")
#LISSY Code
count <- count(breweries, breweries$State, sort=FALSE)
names(count)[1] <- "State"
names(count)[2] <- "Breweries"
count_wrap <- cbind(count[1:(nrow(count)/5), ], count[(1+(nrow(count)/5)):(10+(nrow(count)/5)), ], count[(11+(nrow(count)/5)):(20+(nrow(count)/5)),], count[(21+(nrow(count)/5)):(30+(nrow(count)/5)),], count[(31+(nrow(count)/5)):(40+(nrow(count)/5)),])
count_wrap 
```

#### 2. Merge beer data with breweries data by brewery id. Print ﬁrst 6 observations and the last six observations to check the merged ﬁle.
 we merge `Beers` and `Brewery` datasets by `Brewery_ID` using merge command for base R and assign the new dataset to `Brewery_and_Beer`. We use `head()` and `tail()` to print the first and last 6 rows of the newly created `Brewery_and_Beer` dataset respectively. 
```{r}

#Andy Code
# merge two data frames by ID
names(beers)[5]<- "Brew_ID" #making the merged columns the same
Brewery_and_Beer <- merge(x=beers, y=breweries, by="Brew_ID")

print(head(Brewery_and_Beer)) # printing the first 6 observations

print(tail(Brewery_and_Beer)) #printing the last 6 observations

#LISSY Code
names(beers)[5]<- "Brew_ID" #making the merged columns the same
Breweries_and_Beer <- merge(beers, breweries, by="Brew_ID")
names(Breweries_and_Beer)[2] <- "BeerName" #changing name.x to BeerName
names(Breweries_and_Beer)[8] <- "BreweryName" #changing name.y to BreweryName
head(Breweries_and_Beer, 6)
tail(Breweries_and_Beer, 6)


```

#### 3. Report the number of NA’s in each column.
 as shown in the code block below returns the summary of the number of NA's per column. International Bitterness Units of beer (IBU) has the highest number of NA's  of all the available variables which is 1005.
```{r}
#Lissy Code
for (i in 1:10){
  print(paste(names(Breweries_and_Beer)[i],":", sum(is.na(Breweries_and_Beer[,i]))))}

```

#### 4. Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.
This code block then computes the median alcohol content(ABV) per state by calling `ABV_by_State.R` and stores the result in a dataframe `ABV_by_State`. It also computes median International Bitterness Units of the beer (IBU) per State by calling  `IBU_by_State.R` and stores the result in `IBU_by_State`. `Bar_chart_Plotter.R`, plots a grid bar charts to comparing median ABV and median IBU in each of the 51 States. 
```{r}

# Andy Code
abv <- tapply(Breweries_and_Beer$ABV, Breweries_and_Beer$State, FUN=median, na.rm=TRUE)
abv

print("Median ABV by State")

ibu <- tapply(Breweries_and_Beer$IBU, Breweries_and_Beer$State, FUN=median, na.rm=TRUE)
ibu
print("Median IBU by State")
print("Plotting Bar Chart")

#Lissy Code
abv <- tapply(Breweries_and_Beer$ABV, Breweries_and_Beer$State, FUN=median, na.rm=TRUE)
abv

ibu <- tapply(Breweries_and_Beer$IBU, Breweries_and_Beer$State, FUN=median, na.rm=TRUE)
ibu

states <- count[,1]
abv_percent <- abv*100 #making these values percents so that the comparisons are easier to see on the graph
head(abv_percent)
medians <- data.frame(ibu,abv_percent)
#medians
ibu_abv <- data.frame(c(medians$ibu,medians$abv),states)
names(ibu_abv)[1] <- "Medians"
ibu_abv$Measure <- c(rep("IBU",length(ibu)),rep("ABV",length(abv)))
ibu_abv
ggplot(ibu_abv,aes(State,Medians)) + geom_bar(aes(State,Medians, fill=Measure),stat="identity",position="dodge",width=.7)+scale_fill_manual("Medians\n", values=c("red","blue"), labels=c("ABV","IBU")) + labs(x="\nStates",y="Medians\n")+ theme_bw(base_size=14) + theme(axis.text.x = element_text(angle=90,hjust=1,size=7)) + ggtitle("Comparing Medians of ABV and IBU") 

```


#### 5. Which state has the maximum alcoholic beer? Which state has the most bitter beer?
In this code block, we identify Kentucky(KY) as the State with the maximum alcoholic beer with an `ABV` of *0.125* and Oregon (OR) as the state with the most bitter beer with an `IBU` of *138*.
```{r}
#Andy Code
#maximum alcoholic beer
# select the row with max ABV
Max_ABV<-data.frame(Brewery_and_Beer_Clean[which(Brewery_and_Beer_Clean$ABV==max(Brewery_and_Beer_Clean$ABV)),])

print(paste0("The state with the beer with maximum alcohol is ->", Max_ABV$State, " with an ABV of ", Max_ABV$ABV)) # print the stae with Max alcoholic beer 

#state with the most bitter beer
Max_IBU<-data.frame(Brewery_and_Beer_Clean[which(Brewery_and_Beer_Clean$IBU==max(Brewery_and_Beer_Clean$IBU)),]) # select the row with max IBU

print(paste0("The state with Most bitter beer is ->", Max_IBU$State, " with IBU of ", Max_IBU$IBU)) # print the stae with Max alcoholic beer 

#Lissy Code
abv_clean <- na.omit(Breweries_and_Beer$ABV)
max(abv_clean) #gives us the max ABV, which is 0.128
Breweries_and_Beer[grep("0.128", Breweries_and_Beer$ABV),] #finds the row with the highest alcohol content (.128)

ibu_clean <- na.omit(Breweries_and_Beer$IBU)
max(ibu_clean) #gives us the maximum bitterness, which is 138
Breweries_and_Beer[grep("138", Breweries_and_Beer$IBU),] #finds the row with the highest bitterness (138)

```


#### 6. Summary statistics for ABV (Alcohol by volume) variable.
To get the summary statistics of ABV by Volume variable, I am using describeBy function from pysch package to generate descriptive statistics shown below. To make the summary statistics more readable, I rename the variable column to Volume.
```{r}
#Andy Code
summary(abv_clean)

#Lissy Code
summary(abv_clean)

```

#### 7. Is there a relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.

There is a positive correlation between `ABV` and `IBU` as shown in the regression trend line in the scatter plot below. `IBU` increases with an increase in `ABV`.

```{r}
p <- ggplot(Brewery_and_Beer, aes(Brewery_and_Beer$IBU, Brewery_and_Beer$ABV))
p + geom_point(colour = "red", na.rm=TRUE)+    geom_smooth(method=lm,   # Add linear regression line
                se=FALSE, na.rm=TRUE) +   # Don't add shaded confidence region
  labs(title = "Bitterness of Beer (IBU) vs Alcoholic Content (ABV)", x = "IBU", y ="ABV") +theme(plot.title = element_text(hjust = 0.5))

plot(abv, ibu, main="ABV vs. IBU")
```
