---
title: "Brewery Information"
author: "Lizzy Sterling and Andrew Walch"
date: "10/10/2017"
output: html_document
---
=======
#Introduction#  
The purpose of this code is to take data regarding beer and breweries in the United States of America.  
Within this code, you can find information such as:  
* The number of breweries per state  
* What data is missing and why this matters in the context of this data  
* Bitterness and Alcohol content found in each state, along with comparisons

Libraries used:  
library(dplyr) for the `count` function  
library(ggplot2) for plotting barplot (`ggplot`)
```{r, echo=TRUE, include=FALSE}
library(dplyr) #count function
#bar plot
library(ggplot2)
```
##Reading in the files containing the Beer and the Breweries data in order to conduct further analysis
```{r, echo=TRUE}
knitr::opts_knit$set(root.dir = '/Users/lizzysterling/Desktop/CaseStudy1')
```
`beers` contains information regarding the beer data  
`breweries` contains information regardin the breweries data
>>>>>>> a4d2b5d70c6a201b29d485b6bbb5263c12be2de8
```{r, echo=TRUE}
beers <- read.csv("Beers.csv", header=TRUE)
breweries <- read.csv("Breweries.csv", header=TRUE)
```
##Breweries by State
By using the `count` function within the `dplyr` package, we are able to count the number of breweries found in each separate state.  
The state with the most breweries is Colorado (CO) with 47 breweries  
The states tied with the least breweries are Washington DC, North Dakota (ND), South Dakota (SD), and West Virginia (WV) with 1 brewery
```{r, echo=TRUE}
count <- count(breweries, breweries$State, sort=FALSE)
names(count)[1] <- "State"
names(count)[2] <- "Breweries"
count_wrap <- cbind(count[1:(nrow(count)/5), ], count[(1+(nrow(count)/5)):(10+(nrow(count)/5)), ], count[(11+(nrow(count)/5)):(20+(nrow(count)/5)),], count[(21+(nrow(count)/5)):(30+(nrow(count)/5)),], count[(31+(nrow(count)/5)):(40+(nrow(count)/5)),])
count_wrap #gives the data in a way that shows all of the states in a wrapped fashion
>>>>>>> a4d2b5d70c6a201b29d485b6bbb5263c12be2de8
```
##Merge beer data with the breweries data.
```{r, echo=TRUE}
names(beers)[5]<- "Brew_ID" #making the merged columns the same
Breweries_and_Beer <- merge(beers, breweries, by="Brew_ID")
names(Breweries_and_Beer)[2] <- "BeerName" #changing name.x to BeerName
names(Breweries_and_Beer)[8] <- "BreweryName" #changing name.y to BreweryName
head(Breweries_and_Beer, 6)
tail(Breweries_and_Beer, 6)
>>>>>>> a4d2b5d70c6a201b29d485b6bbb5263c12be2de8
```
#Report the number of NAs in each column
```{r, echo=TRUE}
for (i in 1:10){
  print(paste(names(Breweries_and_Beer)[i],":", sum(is.na(Breweries_and_Beer[,i]))))}
```
##Compute the median alcohol content and international bitterness unit for each state.#
###Median Alcohol Content (ABV)
Utah has the least median alcohol content (4%) - this can most likely be attributed to the large population of Mormons in Utah, and Mormons do not typically drink alcohol, especially when it has a higher alcohol content  
Kentucky and Washington DC are tied for having the highest median alcohol content (6.25%)
```{r, echo=TRUE}
abv <- tapply(Breweries_and_Beer$ABV, Breweries_and_Beer$State, FUN=median, na.rm=TRUE)
abv
```
###Median International Bitterness Unit (IBU)
Maine (ME) has the highest median of bitterness in their beer of 61  
Wisconsin (WI) has the lowest median of bitterness in their beer of 19  
*It is interesting to note that South Dakota apparently has no beer with any bitterness*
```{r, echo=TRUE}
ibu <- tapply(Breweries_and_Beer$IBU, Breweries_and_Beer$State, FUN=median, na.rm=TRUE)
ibu
```
####Plot the Differences
The barplot compares, side by side, IBU and ABV medians. When looking at ABV scores, remember that they are in percentages, while IBU is on a scale from 0-120 (roughly). It is apparent that there is greater range for IBUs, but this may be due to the fact that there is also a wide range of taste preferences when it comes to bitterness. 
```{r, echo=TRUE}
states <- count[,1]
abv_percent <- abv*100 #making these values percents so that the comparisons are easier to see on the graph
head(abv_percent)
medians <- data.frame(ibu,abv_percent)
#medians
ibu_abv <- data.frame(c(medians$ibu,medians$abv),states)
names(ibu_abv)[1] <- "Medians"
ibu_abv$Measure <- c(rep("IBU",length(ibu)),rep("ABV",length(abv)))
ibu_abv
ggplot(ibu_abv,aes(State,Medians)) + geom_bar(aes(State,Medians, fill=Measure),stat="identity",position="dodge",width=.7)+scale_fill_manual("Medians\n", values=c("red","blue"), labels=c("ABV","IBU")) + labs(x="\nStates",y="Medians\n")+ theme_bw(base_size=14) + theme(axis.text.x = element_text(angle=90,hjust=1,size=7)) + ggtitle("Comparing Medians of ABV and IBU") 
```
##Comparing States
###State with the beer with the maximum alcohol content (ABV)
Colorado has the beer with highest alcohol content (12.8%)
```{r, echo=TRUE}
abv_clean <- na.omit(Breweries_and_Beer$ABV)
max(abv_clean) #gives us the max ABV, which is 0.128
Breweries_and_Beer[grep("0.128", Breweries_and_Beer$ABV),] #finds the row with the highest alcohol content (.128)
```
###State with the beer with the maxium bitterness (IBU)
Oregon has the beer with the highest bitterness (138)
```{r, echo=TRUE}
ibu_clean <- na.omit(Breweries_and_Beer$IBU)
max(ibu_clean) #gives us the maximum bitterness, which is 138
Breweries_and_Beer[grep("138", Breweries_and_Beer$IBU),] #finds the row with the highest bitterness (138)
```
##Summary Statistics for the ABV variable
These summary statistics show that there is a large range of alcohol content, with a minimum of .1% and a maximum of 12.8%, with an average of 5.977%. 
```{r, echo=TRUE}
summary(abv_clean)
```
##Relationship Between IBU and ABV of beer
The scatterplot comparing bitterness and alcohol content shows a slight pattern of the higher the alcohol content the higher the bitterness, but there are also many outliers.  
The trend of linearity does not start until after you get to 5% alcohol level.
```{r, echo=TRUE}
plot(abv, ibu, main="ABV vs. IBU")
```
>>>>>>> a4d2b5d70c6a201b29d485b6bbb5263c12be2de8
